Put Title here

### 2018-11-21 General WLD combinatorics.tex
one tiny typo
edited proof of lemma 3.5

\note{I don't think rank of a vertex (in the sense of number of propagators supported on it) was ever used, so I removed the definition}
%\note{rank probably isn't a good word; do we have a better one?}
%\begin{dfn}\label{vertex rank}
%Let $i$ be a vertex in $W$. We define the {\bf rank} of $i$ to be $|Prop(i)|$, i.e. the number of propagators supported on $i$.
%\end{dfn}

@@

\begin{dfn}\note{right now this doesn't depend on the order of $i$ and $j$, but that might matter later on?}
Let $p$ be a propagator supported on the vertices $(i,i+1,j,j+1)$.  Define the \textit{length} of $p$ to be 
\[\ell(p) = min\big\{|[i+1,j]|,|[j+1,i]|\big\}.\]
\end{dfn}

\begin{rmk}\label{rem:props of length 2 and 3}
The following observations about propagators of short length in $W$ are easily verified:
\begin{itemize}
\item If every vertex in $W$ supports at least one propagator, then $W$ admits at least one propagator of length 2.
\item If $p = (i,i+1,i+3,i+4)$ is a propagator of length 3, then the middle vertex $i+2$ supports at most one propagator.
\end{itemize}
\end{rmk}

\note{Are the next few definitions used outside the proof of Lemma~\ref{lem sian 1}?  If not, then perhaps they could go inside the proof so as not to clutter the document with temporary definitions?}
\begin{dfn}
Define the restriction of $W$ to a region $[i,j]$ to be the subdiagram obtained by forgetting all propagators not wholly supported on vertices in the cyclic interval $[i,j]$.  Denote this diagram by $W_{[i,j]}$.
\end{dfn}
\note{I think there's an issue here about what we mean by subdiagram; in def 1.5 we say a subdiagram is of the form $(P, V(P))$ which is certainly not what we want here; in fact I think we want the subdiagram where the vertex set is unchanged and only the propagator set is restricted (the other possibility would be to restrict the vertex set to $[i, j]$, but then the lengths of propagators can change and in the way we want to use it (see paragraph below) $p$ will become very non-admissible).  I think def 1.5 should be removed and we should say that for a WLD $(P, V)$ subdiagrams are any $(P', V')$ with $P'\subseteq P$, $V'\subseteq V$ and all supporting vertices of elements of $P'$ in $V'$; this seems to me to be the natural notion of subobject.}

If $p$ is a propagator supported on $(i,i+1,j,j+1)$, then we can consider two obvious restrictions of $D$, namely $D_{[i,j+1]}$ and $D_{[j,i+1]}$.  Clearly all other propagators $q\neq p$ appear in exactly one of these subdiagrams, while $p$ appears in both.  In this case, we will say that $p$ \textit{bounds the regions} $[i,j+1]$ and $[j,i+1]$; the ordering of the start and end points allow us to identify one or the other region unambiguously.

@@

\begin{lem}\label{lem sian 1}
  Let $W$ be an admissible WLD in which each vertex supports at least one propagator.  Then exactly one of the following two things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length 3 with a propagator of length 2 inside it.
    \item $W$ admits no propagators of length 3, and one or more of the following things occurs.
      \begin{enumerate}
        \item There is a propagator of length $2$, say $(i, i+1, i+2, i+3)$ such that no other propagators ends on either the edge $(i, i+1)$ or the edge $(i+2, i+3)$.\label{item lonely 2}
        \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+1, i+2, i+3)$, the second is $(j, j+1, j+2, j+3)$, no other propagator ends between $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
        \item There exists a propagator of length exactly 4 with only a propagator of length 2 inside it.\label{item 4 and 2}
      \end{enumerate}
  \end{enumerate}
\end{lem}

\begin{proof}
  Suppose first that $W$ has a propagator of length $3$, $p=(i, i+1, i+3, i+4)$.  By Remark~\ref{rem:props of length 2 and 3} along with the fact that every vertex of $W$ supports at least one propagator, we have that $i+2$ supports exactly one propagator and this propagator must have length 2 by noncrossingness.

@@

@@

Start by choosing a propagator $p_1$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}), and write $(j_1,j_1+1,j_1+2,j_1+3)$ for its support.  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise: since $p_1$ is not in configuration \ref{item lonely 2} then there must be a second propagator starting either on edge $(j_1, j_1+1)$ or $(j_1+2, j_1+3)$.  Flipping the diagram if necessary, assume the former.  Since $p_1$ is not in configuration \ref{item pair of 2s} and every vertex is supported then there must exist a propagator $q_1$ of length $\geq 4$ with one end on edge $(j_1,j_1+1)$ or on edge $(j_1-1, j_1)$ or on edge $(j_1-2, j_1-1)$.  This bounds a region $[i_1,k_1]$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.

Now suppose $q_{r}$ exists by the induction hypothesis and bounds a region $[i_r,k_r]$.  Consider the restricted diagram $W_r:=W_{[i_r,k_r]}$; by the original hypotheses on $W$ every vertex $l \in [i_r+2,k_r-2]$ (which is a non-empty interval since $\ell(q_r) \geq 4$) must have support $\geq 1$, while the same is true for $l \in \{i_r,i_r+1,k_r-1,k_r\}$ because of $q_r$.

By Remark~\ref{rem:props of length 2 and 3}, $W_r$ admits at least one propagator of length 2.  $W_r$ has no propagators of length $3$ since $W$ has none. \note{now the question really is important about what length means in the restricted diagram; as written the restricted diagram has the same vertex set so length is unchanged.}   Let $p_{r+1}$ be a propagator of length 2 in $W_r$; if it forms part of configuration \ref{item lonely 2}, \ref{item pair of 2s}, or \ref{item 4 and 2}
%the configuration \eqref{eq:2 prop config}
then we are done, so assume otherwise.

Note we may replace $q_r$ by any other propagator of length $\geq 4$ in $W_r$ which contains $p_{r+1}$ while still satisfying all the hypotheses, and so, without loss of generality we may assume that $q_r$ has minimal length among propagators of length $\geq 4$ which contain $p_{r+1}$.

There are two cases to consider.
The first case is that neither end of $p_{r+1}$ lies on an edge contained in the possibly empty interval $[i_r+3,k_r-3]$, then since $p_{r+1}$ has length 2, it must be that $[i_r+3,k_r-3]$ is at most a single edge, and so $q_r$ has length $\leq 6$.  Since $p_{r+1}$ has one end before $i_r+3$ and the other after $k_r-3$, the only propagators inside $q_r$ that do not also contain $p_{r+1}$ must have both ends in $[i_r, i_r+3]$ or both ends in $[k_r-3, k_r]$ and hence be of length 2.  Therefore, if $q_r$ contains a propagator $t$ of length $\leq 4$, this propagator must also contain $p_{r+1}$ which is impossible my minimality.  Therefore, if $q_r$ contains another propagator of length $2$ we would have configuration \ref{item pair of 2s} which we have already assumed does not occur.  Consequently, $q_r$ contains only $p_{r+1}$; if $q_r$ is of length $4$ then we have \ref{item 4 and 2} and if $q_r$ is of length $5$ or $6$ then we have either \ref{item lonely 2} or an unsupported vertex, all of which are impossible.  Therefore the first case cannot occur.

The second case is that at least one endpoint of $p_{r+1}$ lies on an edge contained in the interval $[i_r+3,k_r-3]$.  Up to symmetry we can write $p_{r+1} = (j_{r+1}, j_{r+1}+1, j_{r+1}+2, j_{r+1}+3)$ where $(j_{r+1}+2, j_{r+1}+3)$ is contained in $[i_r+3, k_r-3]$.  Since $p_{r+1}$ is not part of configuration \ref{item pair of 2s} and vertex $j_{r+1}+4$ is supported (and not by $q_r$ since $j_{r+1}+4 \leq k_r-2$), there is some propagator $t$ with one end in $[j_{r+1}+2, j_{r+1}+4]$ and length $\geq 4$.  By noncrossingness $t$ is contained in $q_{r}$.  $t$ does not contain $p_{r+1}$ by minimality and so we may set $q_{r+1} = t$ to continue the induction.

>>>>

%\note{I don't think rank of a vertex (in the sense of number of propagators supported on it) was ever used, so I removed the definition}
%\note{rank probably isn't a good word; do we have a better one?}
%\begin{dfn}\label{vertex rank}
%Let $i$ be a vertex in $W$. We define the {\bf rank} of $i$ to be $|Prop(i)|$, i.e. the number of propagators supported on $i$.
%\end{dfn}

@@

\begin{dfn}\note{right now this doesn't depend on the order of $i$ and $j$, but that might matter later on?}
Let $p$ be a propagator supported on $(i,j)$.  Define the \textit{length} of $p$ to be 
\[\ell(p) = min\big\{|[i+1,j]|,|[j+1,i]|\big\}.\]
\end{dfn}

\begin{rmk}\label{rem:props of length 2 and 3}
The following observations about propagators of short length in $W$ are easily verified:
\begin{itemize}
\item If every vertex in $W$ supports at least one propagator, then $W$ admits at least one propagator of length 2.
\item If $p = (i,i+3)$ is a propagator of length 3, then the middle vertex $i+2$ supports at most one propagator.
\end{itemize}
\end{rmk}

@@

\begin{lem}\label{lem sian 1}
  Let $W$ be an admissible WLD in which each vertex supports at least one propagator.  Then exactly one of the following two things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length 3 with a propagator of length 2 inside it.
    \item $W$ admits no propagators of length 3, and one or more of the following things occurs.
      \begin{enumerate}
        \item There is a propagator of length $2$, say $(i, i+2)$ such that no other propagators ends on either the edge $i$ or the edge $i+2)$.\label{item lonely 2}
        \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator ends between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
        \item There exists a propagator of length 4 with only a propagator of length 2 inside it.\label{item 4 and 2}
      \end{enumerate}
  \end{enumerate}
\end{lem}

\begin{proof}
  Suppose first that $W$ has a propagator of length $3$, $p=(i, i+3)$.  By Remark~\ref{rem:props of length 2 and 3} along with the fact that every vertex of $W$ supports at least one propagator, we have that $i+2$ supports exactly one propagator and this propagator must have length 2 by noncrossingness.

@@

We need a bit of notation:
Define the restriction of $W$ to a region $[i,j]$ to be the subdiagram obtained by forgetting all propagators not wholly supported on vertices in the cyclic interval $[i,j]$; the vertex set remains the same as in $W$.  Denote this diagram by $W_{[i,j]}$.

Note that if $p$ is a propagator supported on $(i,j)$, then we can consider two obvious restrictions of $W$, namely $W_{[i,j+1]}$ and $W_{[j,i+1]}$.  Clearly all other propagators $q\neq p$ appear in exactly one of these subdiagrams, while $p$ appears in both.  In this case, we will say that $p$ \textit{bounds the regions} $[i,j+1]$ and $[j,i+1]$; the ordering of the start and end points allow us to identify one or the other region unambiguously.

With these observations in mind we can return to the proof.

@@

Start by choosing a propagator $p_1$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}), and write $(j_1,j_1+2)$ for its support.  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise: since $p_1$ is not in configuration \ref{item lonely 2} then there must be a second propagator starting either on edge $j_1)$ or $j_1+2$.  Flipping the diagram if necessary, assume the former.  Since $p_1$ is not in configuration \ref{item pair of 2s} and every vertex is supported then there must exist a propagator $q_1$ of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$.  This bounds a region $[i_1,k_1]$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.

Now suppose $q_{r}$ exists by the induction hypothesis and bounds a region $[i_r,k_r]$ not containing any $p_i$ for $i\leq r$.  Consider the restricted diagram $W_r:=W_{[i_r,k_r]}$; by the original hypotheses on $W$ every vertex $l \in [i_r+2,k_r-2]$ (which is a non-empty interval since $\ell(q_r) \geq 4$) must have support $\geq 1$, while the same is true for $l \in \{i_r,i_r+1,k_r-1,k_r\}$ because of $q_r$.

By Remark~\ref{rem:props of length 2 and 3}, $W_r$ admits at least one propagator of length 2.  $W_r$ has no propagators of length $3$ since $W$ has none.  Let $p_{r+1}$ be a propagator of length 2 in $W_r$; if it forms part of configuration \ref{item lonely 2}, \ref{item pair of 2s}, or \ref{item 4 and 2}
%the configuration \eqref{eq:2 prop config}
then we are done, so assume otherwise.

Note we may replace $q_r$ by any other propagator of length $\geq 4$ in $W_r$ which contains $p_{r+1}$ and such a new $q_r$ still satisfies all the hypotheses, and so, without loss of generality we may assume that $q_r$ has minimal length among propagators of length $\geq 4$ which contain $p_{r+1}$.

There are two cases to consider.
The first case is that $p_{r+1}$ has one end before $i_r+3$ and the other after $k_r-3$.  Then since $p_{r+1}$ has length 2, it must be that $[i_r+3,k_r-3]$ is at most a single edge, and so $q_r$ has length $\leq 6$.  The only propagators inside $q_r$ that do not also contain $p_{r+1}$ must have both ends in $[i_r, i_r+3]$ or both ends in $[k_r-3, k_r]$ and hence be of length 2.  Therefore, if $q_r$ contains a propagator $t$ of length $\leq 4$, this propagator must also contain $p_{r+1}$ which is impossible my minimality.  Therefore, if $q_r$ contains another propagator of length $2$ we would have configuration \ref{item pair of 2s} which we have already assumed does not occur.  Consequently, $q_r$ contains only $p_{r+1}$; if $q_r$ is of length $4$ then we have \ref{item 4 and 2} and if $q_r$ is of length $5$ or $6$ then we have either \ref{item lonely 2} or an unsupported vertex, all of which are impossible.  Therefore the first case cannot occur.

The second case is that either at least one endpoint of $p_{r+1}$ lies on an edge contained in the interval $[i_r+3,k_r-3]$, or both ends lie in $[i_r, i_r+3]$ or $[k_r-3, k_r]$; these situations all behave similarly.  Up to symmetry we can write $p_{r+1} = (j_{r+1}, j_{r+1}+2)$ where either the edge $j_{r+1}+2$ is contained in $[i_r+3, k_r-3]$ or $j_{r+1}+2 = i_r+2$.  Since $p_{r+1}$ is not part of configuration \ref{item pair of 2s} and vertex $j_{r+1}+4$ is supported (and not by $q_r$ since $j_{r+1}+4 \leq k_r-2$ in both situations), there is some propagator $t$ with one end in $[j_{r+1}+3, j_{r+1}+5]$ and length $\geq 4$.  By noncrossingness $t$ is contained in $q_{r}$.  $t$ does not contain $p_{r+1}$ by minimality and so we may set $q_{r+1} = t$ to continue the induction.
